\textbf{quicR}:\ An R Library for Streamlined Data Handling of Real-Time Quaking Induced Conversion Assays

### 2025-05-21 man/manuscript/softwareX_submission.tex
Added clarification that time-to-threshold is also known as lag time.
Addressed reviewer comments.

        \begin{enumerate}
            \item Time-to-threshold (TtT): The time required for the fluorescence signal to exceed a predefined threshold \cite{Orru2015}.
            \item Rate of amyloid formation (RAF): A measure of the kinetics of aggregate growth, which provides insight into the relative quantity of misfolded seed \cite{Gallups2022}.
            \item Maxpoint ratio (MPR): A ratio-based metric measuring peak normalized fluorescence intensities \cite{Rowden2023}.
            \item Maximum slope (MS): The steepest rate of fluorescence increase, reflecting the most rapid phase of aggregation \cite{Henderson2015}.
        \end{enumerate}

@@

@@

        \subsection{Example Code}
            \# Step 1: Identify the raw file.\\
            file <- "example.xlsx"\\
            \# Step 2: Extract the raw data.\\
            raw <- get\_real(file)[[1]]\\
            \# Step 3: Normalize the data against the background.\\
            normal <- normalize\_RFU(raw, transposed = FALSE)\\
            \# Step 4: Extract the metadata.\\
            meta <- organize\_tables(file) |> convert\_tables()\\
            \# Step 5: Get sample locations.\\
            locations <- get\_sample\_locations(file)\\
            \# Step 6: Create the analyzed data frame.\\
            analyzed <- calculate\_metrics(normal, meta)\\
            \# Step 7: Plot the analyzed data frame.\\
            plot\_metrics(analyzed)\\
            \# Step 8: Plot the plate view.\\
            plate\_view(raw, locations)\\

>>>>

        \begin{enumerate}
            \item Time-to-threshold (TtT): The time required for the fluorescence signal to exceed a predefined threshold (also known as lag time) \cite{Orru2015}. 
            \item Rate of amyloid formation (RAF): A measure of the kinetics of aggregate growth, which provides insight into the relative quantity of misfolded seed \cite{Gallups2022}.
            \item Maxpoint ratio (MPR): A ratio-based metric measuring peak normalized fluorescence intensities \cite{Rowden2023}.
            \item Maximum slope (MS): The steepest rate of fluorescence increase, reflecting the most rapid phase of aggregation \cite{Henderson2015}.
        \end{enumerate}

@@

        \subsection{Comparable Methods}
            To contextualize the functionality and performance of \textbf{quicR}, we compared it with existing RT-QuIC analysis approaches from the published R package, \textbf{QuICSeedR} \cite{Li2025}. While both \textbf{quicR} and \textbf{QuICSeedR} offer overlapping core functionalities, their design philosophies, input handling, and computational trade-offs differ substantially.

            Benchmarking was performed using equivalent functions between the two packages, and results are summarized in Table~\ref{tbl:benchmark}. In general, \textbf{QuICSeedR} out-performs \textbf{quicR} across the board. However, \textbf{quicR} addresses several practical limitations that the authors believe justifies the slower times. Notably, \textbf{QuICSeedR} requires an additional, manually formatted Excel file to input sample metadata, whereas \textbf{quicR} directly parses raw and meta data embedded within a single MARS export file. This design reduces user overhead and minimizes potential for transcription errors.

            \textbf{quicR} also provides more granular control over metric calculations, allowing users to compute MPR, MS, TtT, and RAF independently rather than as a fixed batch. While \textbf{QuICSeedR} includes built-in support for bulk analysis and diagnostic outputs, these features can be approximated in \textbf{quicR} via standard R idioms (e.g., \textbf{apply} family of functions), and are planned for future development.

            Importantly, \textbf{quicR} is actively maintained, in contrast to \textbf{QuICSeedR}, which, to the authors’ knowledge, is no longer under active development. Overall, \textbf{quicR} favors flexibility, reproducibility, and usability with raw instrument output at the expense of execution speed—trade-offs that align with the package’s intended use in diagnostic pipelines.

            \begin{table}[ht]
                \centering
                \begin{tblr}{
                    cells     = {font = \fontsize{11pt}{11pt}\selectfont},
                    colspec   = {rcc}, 
                    row{1}    = {font=\bfseries},
                    rowhead   = 1,
                    width     = \textwidth
                }
                    \hline
                    \SetCell[c=1]{c}Feature / Benchmark & \textbf{quicR} & \textbf{QuICSeedR} \\ 
                    \hline
                    Number of input files required & 1 & 2 \\ 
                    Number of threshold methods & 2 & 3 \\
                    Requires manual metadata entry & No & Yes \\
                    Actively maintained & Yes & No \\
                    Built-in bulk analysis support & No & Yes \\
                    Diagnostic analysis functionality & No & Yes \\
                    Incorporates dilution factors & Yes & No \\
                    Converts time units & No & Yes \\
                    Calculates individual metrics independently & Yes & No \\
                    \hline
                    Plate view plot execution time (s) & 0.0978 & 0.0616 \\
                    Metric plot execution time (s) & 0.0108 & 0.0052 \\
                    Metric calculations (MPR, MS, etc.) (s) & 0.3912 & 0.0184 \\
                    Metadata parsing time (s) & 0.1376 & 0.0374 \\
                    Raw data parsing time (s) & 0.0722 & 0.0498 \\
                    Bulk analysis (6 files) (s) & 10.9738 & 0.8486 \\
                    \hline
                \end{tblr}
                \caption{Comparison of core features and execution times between \textbf{quicR} and \textbf{QuICSeedR}. While \textbf{QuICSeedR} demonstrates superior execution speed, \textbf{quicR} offers enhanced metadata handling, metric flexibility, and support for unprocessed instrument output files. All benchmarks were performed using the same dataset and machine (Windows 11, 16GB RAM).}
                \label{tbl:benchmark}
            \end{table}

        \subsection{Maintenance and Community Involvement}
            The \textbf{quicR} package is developed with long-term maintainability and community engagement in mind. The source code is hosted on GitHub (\url{https://github.com/gagerowden/quicR}), where users can report issues, request features, and contribute improvements via pull requests. A comprehensive \textbf{CONTRIBUTING.md} file and issue templates are provided to streamline collaboration and lower the barrier for external contributions. In addition to continuous integration testing, we plan to tag releases with semantic versioning and maintain a changelog to track updates and improvements. Feedback from diagnostic users has already led to enhancements in metadata parsing and output plotting, and we anticipate continued improvements through engagement with the broader scientific community. We encourage users to share their use cases, raise issues, and help shape future development directions.

@@

        \subsection{Example Code}
            \# Step 1: Install and load the package\\
            install.packages("quicR")\\
            library(quicR)\\
            \# Step 2: Identify the raw file.\\
            file <- "example.xlsx"\\
            \# Step 3: Extract the raw data.\\
            raw <- get\_real(file)[[1]]\\
            \# Step 4: Normalize the data against the background.\\
            normal <- normalize\_RFU(raw, transposed = FALSE)\\
            \# Step 5: Extract the metadata.\\
            meta <- organize\_tables(file) |> convert\_tables()\\
            \# Step 6: Get sample locations.\\
            locations <- get\_sample\_locations(file)\\
            \# Step 7: Create the analyzed data frame.\\
            analyzed <- calculate\_metrics(normal, meta)\\
            \# Step 8: Plot the analyzed data frame.\\
            plot\_metrics(analyzed)\\
            \# Step 9: Plot the plate view.\\
            plate\_view(raw, locations)\\
